---
title: "Assignment 11"
author: "Danica Butler"
date: "October 22, 2018"
output: word_document
---
#Number 1
```{r}
adult <- read.csv("C:/Users/student/Documents/Senior Year/Programming with R/adult.csv", header=FALSE, na.strings = " ?")
colnames(adult) <- c('age', 'workclass', 'fnlwgt', 'educatoin', 'educatoin_num', 'marital_status', 'occupation', 'relationship', 'race', 'sex', 'capital_gain', 'capital_loss', 'hours_per_week', 'native_country', 'income')
```
                  
#Number 2
```{r}
categorical <- !sapply(adult, is.numeric)

Mode <- function(x) { 
      Unique <- sort(unique(x))
      Unique[which.max(tabulate(match(x, Unique)))] 
}

adult[categorical] <- lapply(adult[categorical], function(x)
              replace(x, is.na(x), Mode(x[!is.na(x)])))
```

#Number 3
```{r}
library(caret)
numeric <- sapply(adult, is.numeric)
adult[numeric] <- lapply(adult[numeric], function(x)
              replace(x, is.na(x), median(x[!is.na(x)])))
```

#Number 4
```{r}
levels(adult$native_country) = c("Asia", "Americas", "China", "Americas", "Americas", "Americas", "Americas", "Americas", "Europe", "Europe", "Europe", "Europe", "Americas", "Americas", "Europe", "Americas", "China", "Asia", "Asia", "Asia", "Europe", "Europe", "Americas", "Asia", "Asia", "Americas", "Other", "Americas", "Americas", "Asia", "Europe", "Europe", "Americas", "Europe", "Other", "China", "Asia", "Other", "Americas", "Asia", "Europe")
table(adult$native_country)
levels(adult$educatoin) <- c("High School", "High School", "High School", "Elementary School", "Elementary School", "Elementary School", "High School", "Associates", "Associates", "Bachelors/higher", "Bachelors/higher", "High School", "Bachelors/higher", "Elementary School", "Bachelors/higher", "Some College")
table(adult$educatoin)
levels(adult$occupation) = c("Office", "Other", "Handywork", "Office", "Farmer", "Services", "Handywork", "Services", "Services", "Office", "Other", "Office", "Office", "Handywork")
table(adult$occupation)
```

#Number 5
```{r}
dummies_model <- dummyVars(income ~ ., data=adult)
adultData_mat <- predict(dummies_model, newdata = adult)

adultData <- data.frame(adultData_mat)
adultData$income <- adult$income
```

#Number 6
```{r}
preProcess_missingdata_model <- preProcess(adult, method= c("center", "scale"))
preProcess_missingdata_model
```

#Number 7
```{r}
library(caret)
SplitIndex <- createDataPartition(adult$income, p = .70, list = FALSE, times = 1)
train <- adult[SplitIndex,set.seed(2018)]
test <- adult[-SplitIndex,set.seed(2018)]
```

#Number 8
```{r}
library(ranger)
model = ranger(income ~., data = adult)
pred3  = predict(model, data = adult)$predictions
cm3 = confusionMatrix(pred3, adult$income, positive=" >50K")
print(cm3)
```

#Number 9
```{r}
#3
MissingNumber=function(x){
  for (i in 1:ncol(x)) {
    if (is.numeric(x[,i])) {
      x[,i][is.na(x[,i])]=mean(x[,i], na.rm=TRUE)
    }
  }
  return(x)
}
adult9 = MissingNumber(adult)
sum(is.na(adult9))


levels(adult9$native_country) = c("Asia", "Americas", "China", "Americas", "Americas", "Americas", "Americas", "Americas", "Europe", "Europe", "Europe", "Europe", "Americas", "Americas", "Europe", "Americas", "China", "Asia", "Asia", "Asia", "Europe", "Europe", "Americas", "Asia", "Asia", "Americas", "Other", "Americas", "Americas", "Asia", "Europe", "Europe", "Americas", "Europe", "Other", "China", "Asia", "Other", "Americas", "Asia", "Europe")
table(adult9$native_country)
levels(adult9$educatoin) <- c("High School", "High School", "High School", "Elementary School", "Elementary School", "Elementary School", "High School", "Associates", "Associates", "Bachelors/higher", "Bachelors/higher", "High School", "Bachelors/higher", "Elementary School", "Bachelors/higher", "Some College")
table(adult9$educatoin)
levels(adult9$occupation) = c("Office", "Other", "Handywork", "Office", "Farmer", "Services", "Handywork", "Services", "Services", "Office", "Other", "Office", "Office", "Handywork")
table(adult9$occupation)


#5
dummies_model <- dummyVars(income ~ ., data=adult9)
adultData_mat9 <- predict(dummies_model, newdata = adult9)

adultData <- data.frame(adultData_mat9)
adultData$income <- adult9$income



#6
preProcess_missingdata_model <- preProcess(adult9, method= c("center", "scale"))
preProcess_missingdata_model


#7
splitIndex <- createDataPartition(adult9$income, p = .70, list = FALSE, times = 1)
train <- adult9[ splitIndex,]
test <- adult9[-splitIndex,]


#8
library(ranger)
model = ranger(income ~., data = train)
pred9  = predict(model, data = test)$predictions
cm9 = confusionMatrix(pred9, test$income, positive=" >50K")
print(cm9)

```
#Number 10
```{r}
#3
MissingNumbers2=preProcess(adult, method='knnImpute')
adult10=predict(MissingNumbers2, newdata = adult)
sum(is.na(adult10))


levels(adult10$native_country) = c("Asia", "Americas", "China", "Americas", "Americas", "Americas", "Americas", "Americas", "Europe", "Europe", "Europe", "Europe", "Americas", "Americas", "Europe", "Americas", "China", "Asia", "Asia", "Asia", "Europe", "Europe", "Americas", "Asia", "Asia", "Americas", "Other", "Americas", "Americas", "Asia", "Europe", "Europe", "Americas", "Europe", "Other", "China", "Asia", "Other", "Americas", "Asia", "Europe")
table(adult10$native_country)
levels(adult10$educatoin) <- c("High School", "High School", "High School", "Elementary School", "Elementary School", "Elementary School", "High School", "Associates", "Associates", "Bachelors/higher", "Bachelors/higher", "High School", "Bachelors/higher", "Elementary School", "Bachelors/higher", "Some College")
table(adult10$educatoin)
levels(adult10$occupation) = c("Office", "Other", "Handywork", "Office", "Farmer", "Services", "Handywork", "Services", "Services", "Office", "Other", "Office", "Office", "Handywork")
table(adult10$occupation)

dummies_model <- dummyVars(income ~ ., data=adult10)
adult10Data_mat <- predict(dummies_model, newdata = adult10)

adult10Data <- data.frame(adult10Data_mat)
adult10Data$income <- adult10$income



preProcess_missingdata_model <- preProcess(adult10, method= c("center", "scale"))
preProcess_missingdata_model


splitIndex <- createDataPartition(adult10$income, p = .70, list = FALSE, times = 1)
train <- adult10[ splitIndex,]
test <- adult10[-splitIndex,]


#8
library(ranger)
model = ranger(income ~., data = train)
pred10  = predict(model, data = test)$predictions
cm10 = confusionMatrix(pred10, test$income, positive=" >50K")
print(cm10)

```

#Number 11
```{r}
MissingNumbers3=preProcess(adult, method='medianImpute')
adult11=predict(MissingNumbers3, newdata = adult)
sum(is.na(adult11))


levels(adult11$native_country) = c("Asia", "Americas", "China", "Americas", "Americas", "Americas", "Americas", "Americas", "Europe", "Europe", "Europe", "Europe", "Americas", "Americas", "Europe", "Americas", "China", "Asia", "Asia", "Asia", "Europe", "Europe", "Americas", "Asia", "Asia", "Americas", "Other", "Americas", "Americas", "Asia", "Europe", "Europe", "Americas", "Europe", "Other", "China", "Asia", "Other", "Americas", "Asia", "Europe")
table(adult11$native_country)
levels(adult11$educatoin) <- c("High School", "High School", "High School", "Elementary School", "Elementary School", "Elementary School", "High School", "Associates", "Associates", "Bachelors/higher", "Bachelors/higher", "High School", "Bachelors/higher", "Elementary School", "Bachelors/higher", "Some College")
table(adult11$educatoin)
levels(adult11$occupation) = c("Office", "Other", "Handywork", "Office", "Farmer", "Services", "Handywork", "Services", "Services", "Office", "Other", "Office", "Office", "Handywork")
table(adult11$occupation)

dummies_model <- dummyVars(income ~ ., data=adult11)
adult11Data_mat <- predict(dummies_model, newdata = adult11)

trainData <- data.frame(adult11Data_mat)
trainData$income <- adult11$income


numdf <- c()
catdf <- c()
  for(i in 1:ncol(trainData)) {
    if(is.numeric(trainData[,i])) {
      numdf <- c(numdf, names(trainData)[i])
    }
    else {
      catdf <- c(catdf, names(trainData)[i])
    }
  }
  numdata <- trainData[numdf]
  catdata <- trainData[catdf]


trainData2 <- preProcess(numdata, method= c("center", "scale"))
trainData2



names(adult11)[15] = "income"
set.seed(2018)
splitIndex <- createDataPartition(adult11$income, p = .70, list = FALSE, times = 1)
train <- adult11[ splitIndex,]
test <- adult11[-splitIndex,]


library(ranger)
model = ranger(income ~., data = train)
pred11  = predict(model, data = test)$predictions
cm11 = confusionMatrix(pred11, test$income, positive=" >50K")
print(cm11)

```
#Number 12
```{r}

adult$workclass = as.numeric(adult$workclass)
adult$educatoin = as.numeric(adult$educatoin)
adult$marital_status = as.numeric(adult$marital_status)
adult$occupation = as.numeric(adult$occupation)
adult$relationship = as.numeric(adult$relationship)
adult$race = as.numeric(adult$race)
adult$sex = as.numeric(adult$sex)
adult$native_country = as.numeric(adult$native_country)

#1
adult[adult==' ?']=NA
sum(is.na(adult))

#2
missingval=function(x) {
  for(i in 1:ncol(x)){
    if(!is.numeric(x[,i])){
      levels=unique(x[,i])
      x[,i][is.na(x[,i])]=levels[which.max(tabulate(match(x, levels)))]
    }
  }
  return(x)
}
adult2=missingval(adult)
sum(is.na(adult2))

#3
MissingNumbers4=preProcess(adult, method='medianImpute')
adult3=predict(MissingNumbers4, newdata = adult)
sum(is.na(adult3))

#5

dummies_model <- dummyVars(income ~ ., data=adult)
adultData_mat <- predict(dummies_model, newdata = adult)

adultData <- data.frame(adultData_mat)
adultData$income <- adult$income

#6
preProcess_missingdata_model <- preProcess(adult3, method= c("center", "scale"))
preProcess_missingdata_model


#7
splitIndex <- createDataPartition(adult3$income, p = .70, list = FALSE, times = 1)
train <- adult3[ splitIndex,]
test <- adult3[-splitIndex,]


#8
library(ranger)
model = ranger(income ~., data = train)
pred12  = predict(model, data = test)$predictions
cm12 = confusionMatrix(pred12, test$income, positive=" >50K")
print(cm12)
```
```


#Number 13
```{r}
adult[adult==' ?']=NA
sum(is.na(adult))

#2
missingval=function(x) {
  for(i in 1:ncol(x)){
    if(!is.numeric(x[,i])){
      levels=unique(x[,i])
      x[,i][is.na(x[,i])]=levels[which.max(tabulate(match(x, levels)))]
    }
  }
  return(x)
}
adult2=missingval(adult)
sum(is.na(adult2))

#3
num_miss=preProcess(adult2, method='medianImpute')
adult6=predict(num_miss, newdata = adult2)
sum(is.na(adult6))

#5
dummies_model <- dummyVars(income ~ ., data=adult6)
adult6Data_mat <- predict(dummies_model, newdata = adult6)

adult6Data <- data.frame(adult6Data_mat)
adult6Data$salary <- adult6$salary


#6
preProcess_missingdata_model <- preProcess(adult6, method= c("center", "scale"))
preProcess_missingdata_model

#7
splitIndex <- createDataPartition(adult6$income, p = .70, list = FALSE, times = 1)
train <- adult6[ splitIndex,]
test <- adult6[-splitIndex,]


#8
library(ranger)
model = ranger(income ~., data = train)
pred13  = predict(model, data = test)$predictions
cm13 = confusionMatrix(pred13, test$income, positive=" >50K")
print(cm13)
```